### 2019-03-07 referee_response.txt
03/06/2018

Dear Anonymous Referee,

Thank you so much for your kind words and comments on the manuscript "A VLT/FORS2 Narrowband
Imaging Search for MgII Emission in z~0.70 Galaxies". Below is a list of your comments and
our edits according to these comments.

Comment:"In the last paragraph of Section 2.1 a little more information about the absorption lines would be helpful. 
The velocities are quoted as ~150-420 km/s -- is this measured from the centroids of the lines? 
As the authors know, the velocity of MgII can be significantly affected by infilling from emission, so a bit 
more detail on where these numbers come from and how to interpret them would be useful. 
It would also be helpful to note that the equivalent widths given here are in the rest frame, since 
they are discussed in the observed frame in most of the rest of the paper."

Reply: We expanded the last paragraph to include a refined description of the velocities being reported. We 
added the velocity reported for each galaxy in Table 1. Additionally, we discussed what type of model is being used
to measure the lines and how these mitigate the effect of infilling emission. We also note that the EW's mentioned
here are the rest-frame equivalent widths.

Comment: Please add a note to Table 1 explaining what tau_V is, since it isn't discussed until later in the paper
and it is a bit less obvious than the other quantities in the table.

Reply: We added a note to Table 1 which gives more insight to tau_V.

Comment:In the discussion of previous results in Section 6.1, I think providing some context from Henry et al 2018 would
be useful. Although this paper didn't present spatially extended emission, their results on which galaxies are 
likely to have strong MgII emission is relevant, in particular the finding that nebular emission is important as 
well as resonantly scattered continuum photons, and the finding that the strength of the emission may be set by the 
escape fraction of MgII photons rather than the intrinsic production.

Reply: Added a discussion of Henry et al 2016 and Guseva et al.(2019).