### 2019-03-27 referee_response.txt
Kate's final edits

We wish to thank the referee for their supportive and helpful comments on our
manuscript "A VLT/FORS2 Narrowband Imaging Search for MgII Emission in
z~0.70 Galaxies".  We have implemented each of the suggestions they
sent in their report (in boldface type).  Please see our direct
responses to each comment below.

Comment:"In the last paragraph of Section 2.1 a little more information about the absorption lines would be helpful. 
The velocities are quoted as ~150-420 km/s -- is this measured from the centroids of the lines? 
As the authors know, the velocity of MgII can be significantly affected by infilling from emission, so a bit 
more detail on where these numbers come from and how to interpret them would be useful. 
It would also be helpful to note that the equivalent widths given here are in the rest frame, since 
they are discussed in the observed frame in most of the rest of the paper."

Reply: We expanded the last paragraph of Section 2.1 to include a more
detailed description of the velocities being reported.  We also
added the velocity reported for each galaxy to Table 1. Additionally,
we now discuss the type of model that was used
to fit the lines and how this mitigates the effects of infilling from scattered
emission.  We have added a note stating that the EWs mentioned
here are rest-frame equivalent widths.

Comment: "Please add a note to Table 1 explaining what tau_V is, since it isn't discussed until later in the paper
and it is a bit less obvious than the other quantities in the table."

Reply: We have added a note to Table 1 which gives some description of tau_V.

Comment: "In the discussion of previous results in Section 6.1, I think providing some context from Henry et al 2018 would
be useful. Although this paper didn't present spatially extended emission, their results on which galaxies are 
likely to have strong MgII emission is relevant, in particular the finding that nebular emission is important as 
well as resonantly scattered continuum photons, and the finding that the strength of the emission may be set by the 
escape fraction of MgII photons rather than the intrinsic production."

Reply: We have added discussion of results from Henry et al (2018) to
a new section following Section 6.1 as requested (see Section 6.2).